LC-MS ^13^C Lactate Measurements

### 2022-12-19 analysis/glc6-lactate.qmd
work on cleaning pipeline for glc6 fluxes

withr::with_dir(here::here(), {
  path <- tar_read(fluxes_glc6_files)
})
```

@@

```{r}
#| label: raw-data
df <-
  read_ms_excel(here::here(path)) |> 
  dplyr::filter(!str_detect(.data$id, "B")) |> 
  tidyr::separate(
    .data$sample, 
    c("type", "number"), 
    sep = "(?<=[[:alpha:]])(?=[[:digit:]])", 
    fill = "left", 
    convert = TRUE
  ) |>
  dplyr::mutate(
    conc = ifelse(is.na(.data$type), .data$number, NA_character_), 
    date = dplyr::case_when(
      .data$type == "A" ~ "2022-11-12", 
      .data$type == "B" ~ "2022-11-17", 
      .data$type == "C" ~ "2022-11-22"
    ), 
    number = ifelse(!is.na(conc), NA, .data$number), 
    batch = "b", 
    cell_type = "lf", 
    experiment = "substrate", 
    type = dplyr::case_when(
      .data$type == "QC" ~ "qc", 
      is.na(.data$type) ~ "std", 
      TRUE ~ "sample"
    ), 
    detector = "lcms", 
    run = "a"
  ) |> 
  dplyr::select(
    "metabolite", 
    "detector", 
    "mz", 
    "rt", 
    "cell_type",
    "experiment", 
    "batch", 
    "date", 
    "run", 
    "id", 
    "type", 
    "number", 
    "conc", 
    "area"
  )
```

@@

df_drift <- 
  df |> 
  filter(type == "qc") |> 
  group_by(metabolite) |> 
  mutate(norm = area / mean(area))

ggplot(df_drift) +
  aes(
    x = number, 
    y = norm, 
    color = metabolite
  ) +
  geom_line() + 
  labs(
    x = "Run order", 
    y = "Normalized peak area", 
    color = "Metabolite", 
    title = "QC Drift"
  )
```

@@

ggplot(df_drift) +
  aes(
    x = metabolite, 
    y = norm, 
    color = metabolite
  ) +
  geom_boxplot(show.legend = FALSE) +
  geom_jitter(
    width = 0.1, 
    show.legend = FALSE
  ) +
  labs(
    x = "Metabolite", 
    y = "Normalized peak area", 
    title = "QC Variance"
  )
```

@@

```{r}
#| label: se

se <- 
  df |> 
  new_tbl_se(
    a_data = "area", 
    f_names = "metabolite", 
    f_data = c("metabolite", "detector", "mz", "rt"), 
    s_names = "id", 
    s_data = c(
      "cell_type", 
      "experiment", 
      "batch", 
      "run", 
      "date", 
      "type", 
      "number", 
      "conc"
    )
  ) |> 
  tbl_to_se()

se_drift <- 
  correct_drift(se) |> 
  se_to_tbl(rownames = "metabolite")
```

@@

se_drift |> 
  filter(type == "qc") |> 
  group_by(metabolite) |> 
  mutate(norm = area / mean(area)) |> 
  ggplot() +
  aes(
    x = number, 
    y = norm, 
    color = metabolite
  ) +
  geom_line() + 
  labs(
    x = "Run order", 
    y = "Normalized peak area", 
    color = "Metabolite", 
    title = "Drift Corrected"
  )
```

>>>>

withr::with_dir(here::here(), {
  se_raw <- tar_read(fluxes_glc6_se)
  df_drift <- tar_read(fluxes_glc6_drift)
})
```

@@

```{r}
#| label: raw-data
# df <-
#   read_ms_excel(here::here(path)) |> 
#   dplyr::filter(!str_detect(.data$id, "B")) |> 
#   tidyr::separate(
#     .data$sample, 
#     c("type", "number"), 
#     sep = "(?<=[[:alpha:]])(?=[[:digit:]])", 
#     fill = "left", 
#     convert = TRUE
#   ) |>
#   dplyr::mutate(
#     conc = ifelse(is.na(.data$type), .data$number, NA_character_), 
#     date = dplyr::case_when(
#       .data$type == "A" ~ "2022-11-12", 
#       .data$type == "B" ~ "2022-11-17", 
#       .data$type == "C" ~ "2022-11-22"
#     ), 
#     number = ifelse(!is.na(conc), NA, .data$number), 
#     batch = "b", 
#     cell_type = "lf", 
#     experiment = "substrate", 
#     type = dplyr::case_when(
#       .data$type == "QC" ~ "qc", 
#       is.na(.data$type) ~ "std", 
#       TRUE ~ "sample"
#     ), 
#     detector = "lcms", 
#     run = "a"
#   ) |> 
#   dplyr::select(
#     "metabolite", 
#     "detector", 
#     "mz", 
#     "rt", 
#     "cell_type",
#     "experiment", 
#     "batch", 
#     "date", 
#     "run", 
#     "id", 
#     "type", 
#     "number", 
#     "conc", 
#     "area"
#   )
```

@@

df_raw <- 
  se_raw |> 
  se_to_tbl(rownames = "metabolite") |> 
  filter(type == "qc") |> 
  group_by(metabolite) |> 
  mutate(norm = area / mean(area))

ggplot(df_raw) +
  aes(
    x = number, 
    y = norm, 
    color = metabolite
  ) +
  geom_line() + 
  labs(
    x = "Run order", 
    y = "Normalized peak area", 
    color = "Metabolite", 
    title = "QC Drift"
  )
```

@@

ggplot(df_raw) +
  aes(
    x = metabolite, 
    y = norm, 
    color = metabolite
  ) +
  geom_boxplot(show.legend = FALSE) +
  geom_jitter(
    width = 0.1, 
    show.legend = FALSE
  ) +
  labs(
    x = "Metabolite", 
    y = "Normalized peak area", 
    title = "QC Variance"
  )
```

@@

@@

df_drift |> 
  filter(type == "qc") |> 
  group_by(metabolite) |> 
  mutate(norm = area / mean(area)) |> 
  ggplot() +
  aes(
    x = number, 
    y = norm, 
    color = metabolite
  ) +
  geom_line() + 
  labs(
    x = "Run order", 
    y = "Normalized peak area", 
    color = "Metabolite", 
    title = "Drift Corrected"
  )
```
